A Formal Model of Checked C

### 2022-08-09 sandp23/introduction.tex
New intro

 and there has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area. Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning. E.g., Given sensitive data in a program, the goal is to partition functions into two parts or partitions based on whether a function can access the sensitive data.
The performance overhead of these approaches is dominated by marshaling costs and depends on the usage of sensitive data.
The overhead of state-of-the-art approaches~\cite{lind2017glamdring, liu2017ptrsplit} is prohibitive and varies from 37\%-163\%.
But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code. We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.

The guarantee \checkedc provides focuses on checked code regions,
where any stuck (i.e., ill-defined)
state reached by a well-typed program amounts to a spatial safety
violation.
Such a state can always be attributed to, i.e., 
\emph{blamed on}, the execution of code that is not in a checked core region. 
However, what are the guarantees that unchecked (unsafe) code regions can provide?
Programmers might not expect to convert every code to \checkedc,
but leave some functions in unsafe regions,
in which case the unsafe region executions should not affect other parts.
In this paper, we cover this gap by introducing \systemname,
combining \checkedc with program partitioning mechanism, 
making three main contributions.

\myparagraph{Executions Cannot Crash}
Our first contribution is a clear program partition between checked and unchecked code regions
and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
We restrict the checked and unchecked pointers usage only in checked and unchecked code regions, respectively;
and develop tainted pointers for the communication between the two regions.
Heap is partitioned into two parts: a \textit{checked heap} containing all checked pointers
 and an \textit{unchecked heap} that is sandboxed.
Tainted pointers exist in the unchecked heap region and must be verified each time being used.

>>>>

In this work, we propose a type-directed code-centric program partitioning approach.
Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.

The developer starts by marking desired (i.e., unchecked) functions and pointers used in these functions as tainted.
Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
\begin{itemize}
\item~\ucregion or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (i.e., functions and pointers) and can only access tainted pointers.
\item~\cregion or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
The functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
\end{itemize}
Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (e.g., WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.

The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
We avoid marshaling by allocating all tainted pointers (i.e., tainted buffers) in~\ucregion and thus can be accessed in both partitions.
Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
Our compiler avoids these attacks by ensuring, either statically or through dynamic checks that tainted pointers can only point to~\ucregion address space.

In summary, we make the following three main contributions.

\myparagraph{Type-Directed Program Partitioning}
Our first contribution is a clear program partition between checked and unchecked code regions
and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
We restrict the checked and unchecked pointers usage only in checked and unchecked code regions, respectively;
and develop tainted pointers for the communication between the two regions.
Heap is partitioned into two parts: a \textit{checked heap} containing all checked pointers
 and an \textit{unchecked heap} that is sandboxed.
Tainted pointers exist in the unchecked heap region and must be verified each time being used.
